### 2019-09-12 2-Chapters/4-Chapter/CrownCom_17/IoT_slotted_appendix.tex
Removed raw inclusion of https://hal.inria.fr/hal-02006825 https://hal.inria.fr/hal-02049824v2 https://hal.archives-ouvertes.fr/hal-01575419v2

We include here the missing details of the proof of Proposition~\ref{prop:41:Lagrangian}.
%
First, we need to justify that the objective function is quasi-convex, in each of its coordinates.
%
Then, we develop the computation of $D_i^*(\lambda)$, as a closed form expression of the system parameters ($K, p$), the repartition of static devices ($S_1,\dots,S_{K}$) and the Lagrange multiplier $\lambda$.

>>>>

We include here the missing details of the proof of Proposition~\ref{prop:41:Lagrangian}.
%
First, we need to justify that the objective function is quasi-convex, in each of its coordinates.
%
Then, we develop the computation of $D_i^*(\lambda)$, as a closed form expression of the system parameters ($K, p$), the distribution of static devices ($S_1,\dots,S_{K}$) and the Lagrange multiplier $\lambda$.

### 2019-09-12 2-Chapters/4-Chapter/chapter4.tex
Almost done proofreading résumé fr
OH LA BOULETTE : en anglais, « repartition » n'existe pas. On dit « distribution » ou « allocation »
- https://www.wordreference.com/enfr/repartition
- https://www.deepl.com/translator#en/fr/repartition
Removed raw inclusion of https://hal.inria.fr/hal-02006825 https://hal.inria.fr/hal-02049824v2 https://hal.archives-ouvertes.fr/hal-01575419v2

\abstractStartChapter{}%
%
After detailing in Part~\ref{part:Introduction} the MAB model, we now come back to our main question of interest.
In this chapter, we focus on wireless networks following the hypotheses common to present and future Internet of Things (IoT) networks exposed in Chapter~\ref{chapter:1}.
Our goal is to show that the IoT devices can automatically learn to increase their battery life and their successful transmission rates, without changing anything on the IoT standard side.
%
We propose two models of IoT networks, composed of many independent IoT end-devices, that can use low-cost Reinforcement Learning (RL) algorithms in order to learn how to improve their spectrum access.
%
Decentralized RL for IoT lets the devices use acknowledgements sent back by their Base Station as a reward, instead of sensing feedback like for OSA.
We consider many independent ``dynamic'' devices, each communicating with a small probability at every instant.
Simulations show that dynamic devices can greatly improve their spectrum efficiency,
by using MAB algorithms like \UCB.
% in the case of a non-uniform repartition of the static devices in the channels
We also developed a proof-of-concept using USRP platforms, for a real-world validation of this approach.
% Numerical simulations justify the interest of the approach,
%  of the advantage of using MAB algorithms for the first model.
%
% On a simple example, we present in Section~\ref{sec:4:gnuradio} a demonstration that implements and validates the proposed approach, using real wireless radio hardware.
% The first model
% % without retransmission of packets,
% is interesting for its simplicity,
% and because considering no retransmission can improve the battery life of IoT devices, at the cost of a lower Quality of Service (QoS).
%
In a second step, we consider a second model where a dynamic device has to (try to) retransmit a message, in case of a failed first transmission, up-to a fixed number of retransmissions.
We compare heuristics based on \UCB, and simulations confirm that the non-naive heuristics also significantly improve the network efficiency.
% Interestingly, the simplest heuristic is found to be the most efficient one, as its simplicity implies a faster learning and convergence time.
%
% The second model is more interesting if the main target is an improvement of the QoS, rather than a longer battery life.

@@

After the first chapters that presented the model of MAB, we go back to the initial problems studied in this thesis, and thus we focus on Internet of Things (IoT) networks.
%
As explained in the introduction in Chapter~\ref{chapter:1},
unlicensed bands are more and more used and considered for mobile and LAN (Local Area Network) and for Internet of Things communication standards.
%  (WiFi, LTE-U), (IoT) standards for short-range (ZigBee, Z-Wave, Bluetooth) and long-range (LoRaWAN, SIGFOX, Ingenu, Weightless) communications \cite{Centenaro16}.
This heavy use of unlicensed bands, in particular with the expected exponential growth of the number of IoT devices, will cause performance drop, due to radio collisions that could even compromise IoT promises.

@@

% As detailed in Chapter~\ref{chapter:1},
% MAB learning has already been proposed in Cognitive Radio (CR) \cite{Mitola99,Haykin05}, and in particular, for sensing-based Dynamic Spectrum Access (DSA) in licensed bands \cite{Jouini10}.
% For example,
% proof-of-concepts like \cite{kumar2016two} have proven the capability of such approaches on real radio signals for OSA,
% and \cite{Maghsudi16} shows how MAB learning can be applied for small cell management in licensed 5G networks.
% Some analysis on real radio measurements made for HF ionospheric channels have also proven that solutions based on MAB learning is appropriate and solves efficiently this kind of decision-making problems on real-world wireless signals \cite{Melian15}.
% Recent works show that stationary MAB algorithms work well to solve reinforcement learning models that represent accurately real-world radio problems.
% Recently, TS and \UCB{} algorithms have been used for improving the spectrum access in (unlicensed) WiFi networks, for instance by \cite{Toldov16} or \cite{Wilhelmi19collaborative,Wilhelmi19potential}.
% % However, even with only one dynamic user using the learning algorithm, the background traffic or the traffic of the other devices is never really stationary or \iid{}.

@@

\input{2-Chapters/4-Chapter/CrownCom_17.git/IoT_slotted.tex}

@@

\input{2-Chapters/4-Chapter/IEEE_WCNC_2019__DemoICT.git/main.tex}

@@

\input{2-Chapters/4-Chapter/IEEE_WCNC__2019__Paper__BMBM.git/IEEE_WCNC__2019__Paper__BMBM.tex}

@@

\input{2-Chapters/4-Chapter/CrownCom_17.git/IoT_slotted_appendix.tex}

@@

\input{2-Chapters/4-Chapter/IEEE_WCNC_2019__DemoICT.git/main_appendix.tex}

>>>>

\abstractStartChapter{}%
%
After detailing in Part~\ref{part:Introduction} the MAB model, we now come back to our main question of interest.
In this chapter, we focus on wireless networks following the hypotheses common to present and future Internet of Things (IoT) networks exposed in Chapter~\ref{chapter:1}.
Our goal is to show that the IoT devices can automatically learn to increase their battery life and their successful transmission rates, without changing anything on the IoT standard side.
%
We propose two models of IoT networks, composed of many independent IoT end-devices, that can use low-cost Reinforcement Learning (RL) algorithms in order to learn how to improve their spectrum access.
%
Decentralized RL for IoT lets the devices use acknowledgements sent back by their Base Station as a reward, instead of sensing feedback like for OSA.
We consider many independent ``dynamic'' devices, each communicating with a small probability at every instant.
Simulations show that dynamic devices can greatly improve their spectrum efficiency,
by using MAB algorithms like \UCB.
% in the case of a non-uniform distribution of the static devices in the channels
We also developed a proof-of-concept using USRP platforms, for a real-world validation of this approach.
% Numerical simulations justify the interest of the approach,
%  of the advantage of using MAB algorithms for the first model.
%
% On a simple example, we present in Section~\ref{sec:4:gnuradio} a demonstration that implements and validates the proposed approach, using real wireless radio hardware.
% The first model
% % without retransmission of packets,
% is interesting for its simplicity,
% and because considering no retransmission can improve the battery life of IoT devices, at the cost of a lower Quality of Service (QoS).
%
In a second step, we consider a second model where a dynamic device has to (try to) retransmit a message, in case of a failed first transmission, up-to a fixed number of retransmissions.
We compare heuristics based on \UCB, and simulations confirm that the non-naive heuristics also significantly improve the network efficiency.
% Interestingly, the simplest heuristic is found to be the most efficient one, as its simplicity implies a faster learning and convergence time.
%
% The second model is more interesting if the main target is an improvement of the QoS, rather than a longer battery life.

@@

After the first chapters that presented the model of MAB, we go back to the initial problems studied in this thesis, and thus we focus on Internet of Things (IoT) networks.
%
As explained in the introduction in Chapter~\ref{chapter:1},
unlicensed bands are more and more used and considered for mobile and LAN (Local Area Network) and for Internet of Things communication standards.
%  (Wi-Fi, LTE-U), (IoT) standards for short-range (ZigBee, Z-Wave, Bluetooth) and long-range (LoRaWAN, SIGFOX, Ingenu, Weightless) communications \cite{Centenaro16}.
This heavy use of unlicensed bands, in particular with the expected exponential growth of the number of IoT devices, will cause performance drop, due to radio collisions that could even compromise IoT promises.

@@

% As detailed in Chapter~\ref{chapter:1},
% MAB learning has already been proposed in Cognitive Radio (CR) \cite{Mitola99,Haykin05}, and in particular, for sensing-based Dynamic Spectrum Access (DSA) in licensed bands \cite{Jouini10}.
% For example,
% proof-of-concepts like \cite{kumar2016two} have proven the capability of such approaches on real radio signals for OSA,
% and \cite{Maghsudi16} shows how MAB learning can be applied for small cell management in licensed 5G networks.
% Some analysis on real radio measurements made for HF ionospheric channels have also proven that solutions based on MAB learning is appropriate and solves efficiently this kind of decision-making problems on real-world wireless signals \cite{Melian15}.
% Recent works show that stationary MAB algorithms work well to solve reinforcement learning models that represent accurately real-world radio problems.
% Recently, TS and \UCB{} algorithms have been used for improving the spectrum access in (unlicensed) Wi-Fi networks, for instance by \cite{Toldov16} or \cite{Wilhelmi19collaborative,Wilhelmi19potential}.
% % However, even with only one dynamic user using the learning algorithm, the background traffic or the traffic of the other devices is never really stationary or \iid{}.

@@

\input{2-Chapters/4-Chapter/CrownCom_17/IoT_slotted.tex}

@@

\input{2-Chapters/4-Chapter/IEEE_WCNC_2019__DemoICT/IEEE_WCNC_2019__DemoICT.tex}

@@

\input{2-Chapters/4-Chapter/IEEE_WCNC__2019__Paper__BMBM/IEEE_WCNC__2019__Paper__BMBM.tex}

@@

\input{2-Chapters/4-Chapter/CrownCom_17/IoT_slotted_appendix.tex}

@@

\input{2-Chapters/4-Chapter/IEEE_WCNC_2019__DemoICT/IEEE_WCNC_2019__DemoICT_appendix.tex}
